Variable Selection in Convex Function Estimation

### 2013-11-01 tex/discussion_notes.tex
journal

\begin{document}
\section{(7/26/2013)}

>>>>

\begin{document}\

\section{10/25/2013: Convex + Concave function}

Now we consider variable selection for $h(x) = f(x) + g(x)$ where $f$ is convex and $g$ is concave.

\subsection{Separation Approach}

Here is one line of thought. We are given $h_i = h(x_i)$ for a bunch of sample points. Suppose we can learn the decomposition $f_i + g_i = h_i$, then we can use additive faithfulness to recover the relevant variables of $f$ and $g$.

The separation may not be easy however. Perhaps the following optimization:

\begin{align*}
\min_{\beta, \theta}&\sum_{j=1}^p \Big( \|\beta_j\|_\infty + \| \theta_j\|_\infty \Big) \\
 \trm{s.t. } & h_i = f_i + g_i \\
 & f_i \geq f_{i'} + \theta_{i'}^\tran (x_i - x_{i'}) \\
 & g_i \leq g_{i'} + \beta_{i'}^\tran (x_i - x_{i'}) 
\end{align*}

Of course, we have identifiability issues. There exist two sets of functions $\mathcal{F}$ and $\mathcal{G}$ such that for all $f \in \mathcal{F}$, there exist $g \in \mathcal{g}$ such that $f+g = h$.

The gradient penalty in the optimization program is then critical because it may let us learn a \textbf{sparse} $f,g$.

We emphasize that we do not attempt to do variable selection at all in the separation optimization, we only want to recover the values $f_i$ and $g_i$. The sparsity patterns of $\beta$ and $\theta$ are of no importance.

\section{(7/26/2013)}

### 2013-11-01 tex/scam/additive_faithful.tex
journal

The additive approximation may mark a relevant variable as
irrelevant. Such mistakes are inherent to the approximation and may
persist even with infinite
samples.  Lemma~\ref{lem:general_int_reduction} below provides
examples of errors made by the additive approximations that minimize
the $L_2$ distance.

\begin{lemma}
\label{lem:general_int_reduction}
Let $F$ be a product distribution on $\mathbf{C}=[0,1]^s$ with density function $p$ which is positive on $\mathbf{C}$. Let
$X=(X_1,...,X_s) \sim F$. Let $f: \mathbf{C} \rightarrow \R$ be an
integrable function and suppose, without loss of generality, that $\E f(X) = 0$.
Let $\ds f^*_1,...,f^*_s \coloneqq \arg\min \{ \E \bigl( f(X) -
\sum_{k=1}^s f_k(X_k) \bigr)^2 \,:\, \E f_k(X_k) = 0\}$. Then $f^*_k(x_k) = \E[ f(X) \given x_k]$ is unique.
\end{lemma}

Lemma~\ref{lem:general_int_reduction} follows from the stationarity
condition of the optimal solution. If $F$ is the uniform distribution,
then $f^*_k(x_k) = \int_{\mathbf{x}_{-k}} f(x_k, \mathbf{x}_{-k})
d\mathbf{x}_{-k}$.

\begin{example} We give two examples of additive unfaithfulness under
  the uniform distribution. First, consider the following function:
\[
\trm{(egg carton)} \quad f(x_1, x_2) = \sin( 2\pi x_1) \sin( 2 \pi x_2) \qquad \trm{ for } (x_1, x_2) \in [0,1]^2
\]
For all $x_1$, $\int_{x_2} f(x_1, x_2) d x_2 = 0$ and also, for all
$x_2$, $\int_{x_1} f(x_1, x_2) d x_1 = 0$. An additive approximation
would set $f_1 = 0$ and $f_2 = 0$.  Next, consider the function
\[
\trm{(tilting slope)} \quad f(x_1, x_2) = x_1 x_2 \qquad \trm{ for } x_1 \in [-1,1],\; x_2 \in [0,1].
\]
For all $x_2$, $\int_{x_1} f(x_1, x_2) d x_1 = 0$, therefore, we expect $f_2 = 0$ under the additive approximation. This function, for every fixed $x_2$, is a zero-intercept linear function of $x_1$ with slope exactly $x_2$.
\end{example}

\begin{figure}[htp]
\vskip-10pt
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\centering
		{\includegraphics[width=0.75\textwidth]{figs/sine_wave_funct3}}
		\caption{egg carton}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
		\centering
		{\includegraphics[width=0.75\textwidth]{figs/tilting_slope_funct3}}
		\caption{tilting slope}
	\end{subfigure}

\caption{Two additively unfaithful functions. Relevant variables are
  zeroed out under an additive approximation because every ``slice''
  of the function integrates to zero.}
\vskip-10pt
\end{figure}

It is important to understand the kind of functions for which the
additive approximation can accurately capture all of the relevant variables.
We call this property \emph{additive faithfulness}.

@@

Let $F$ be a probability distribution on $\mathbf{C}$ and
assume without loss of generality that $\E f(X) = 0$. Let 
$$\ds f^*_1,...,f^*_s \coloneqq \arg\min \Bigl\{ \E ( f(X) - \sum_{k=1}^s
f_k(X_k) )^2 \,:\, \E f_k(X_k) = 0, \forall k\Bigr\}.$$
We say that $f$ is \emph{additively faithful} under $F$ in case $f^*_k = 0$ iff $f$ does not depend on coordinate $k$. 
\end{definition}
% We can define the support $\trm{supp}(f) \coloneqq \{ k \,:\,
% \trm{$k$ is relevant to $f$}\}$. Let $f^* = \sum_{k=1}^s$, then $f$
% is additively faith if $\trm{supp}(f) = \trm{supp}(f^*)$.

@@

We give the full proof in Section~\ref{sec:faithful_proof} of the
Appendix, but pause here to provide some intuition. From
Lemma~\ref{lem:general_int_reduction}, we know intuitively that the
additive approximation zeroes out $k$ when, fixing $x_k$, every
``slice'' of $f$ integrates to zero. We prove
Theorem~\ref{thm:convex_faithful} by showing that ``slices'' of convex
functions that integrate to zero cannot be ``glued'' together while
still maintaining convexity.

Theorem~\ref{thm:convex_faithful} plays an important role in our
sparsistency analysis, in which we prove that the additive
approximation is variable selection consistent (or ``sparsistent'') \emph{even when the true function is not
additive.}

\begin{remark}
  We assume twice differentiability in
  Theorem~\ref{thm:convex_faithful} to simplify the proof. We believe
  this smoothness condition is not necessary because every non-smooth
  convex function can be approximated arbitrarily well by a smooth
  one.  Without restrictions on the distribution, a convex
  function may not be additively faithful. Intuitively, an arbitrarily shaped
  density $p$
  may ``undo'' the convexity of $f$ so that the product
  $p(\mathbf{x}) \, f(\mathbf{x})$ resembles an egg carton or a
  tilting slope.  With appropriate conditions on the density $p$,
  however, it is possible to relax the independence assumption.  We leave this to
  future work.
\end{remark}

>>>>

For general regression, additive approximation may result in a
relevant variable being incorrectly marked as irrelevant. Such
mistakes are inherent to the approximation and may persist even with
infinite samples.  In this section we give
examples of this phenomenon, and then show how the convexity
assumption
changes the behavior of the additive approximation.  We begin
with a lemma that implies uniqueness of the additive regression function.

\begin{figure*}[htp]
\vskip-10pt
	\centering
	\subfigure[egg carton]{
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/sine_wave_funct3}}
	}
	\subfigure[tilting slope]{
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/tilting_slope_funct3}}
	}
\caption{Two additively unfaithful functions. Relevant variables are
  zeroed out under an additive approximation because every ``slice''
  of the function integrates to zero.}
\vskip-10pt
\end{figure*}

\begin{lemma}
\label{lem:general_int_reduction}
Let $F$ be a product distribution on $\mathbf{C}=[0,1]^s$ with density function $p$ which is positive on $\mathbf{C}$. Let
$X=(X_1,...,X_s) \sim F$. Let $f: \mathbf{C} \rightarrow \R$ be an
integrable function.
Let 
\begin{equation}
\begin{split}
f_k^*, \mu^* \coloneqq \argmin_{f_1,\ldots, f_s, \mu} \Bigl\{ \E \bigl( f(X) & -
 \sum_{k=1}^s f_k(X_k) -\mu \bigr)^2 \\
       & \,:\, \E f_k(X_k) = 0\Bigr\}.
\end{split}
\end{equation}
Then $\mu^* = \E f(X)$ and $f^*_k(x_k) = \E[ f(X) \given x_k] - \E f(X)$ and this solution is unique.
\end{lemma}

Lemma~\ref{lem:general_int_reduction} follows from the stationarity
conditions of the optimal solution. If $F$ is the uniform distribution,
then $f^*_k(x_k) = \int f(x_k, \mathbf{x}_{-k})
d\mathbf{x}_{-k}$.

\begin{example} We give two examples of additive unfaithfulness under
  the uniform distribution. First, consider the following function:
\[
\trm{(egg carton)} \quad f(x_1, x_2) = \sin( 2\pi x_1) \sin( 2 \pi x_2)
\]
defined for $(x_1, x_2) \in [0,1]^2$.  Then
$\int_{x_2} f(x_1, x_2) d x_2 = 0$ and
$\int_{x_1} f(x_1, x_2) d x_1 = 0$ for each $x_1$ and $x_2$. An additive approximation
would set $f_1 = 0$ and $f_2 = 0$.  Next, consider the function
\[
\trm{(tilting slope)} \quad f(x_1, x_2) = x_1 x_2
\]
defined for $x_1 \in [-1,1],\; x_2 \in [0,1]$.  In this case
$\int_{x_1} f(x_1, x_2) d x_1 = 0$ for each $x_2$; therefore, we expect $f_2 = 0$ under the additive approximation. This function, for every fixed $x_2$, is a zero-intercept linear function of $x_1$ with slope $x_2$.
\end{example}

In order to exploit additive models, it is important to understand when the
additive approximation accurately captures all of the relevant variables.
We call this property \emph{additive faithfulness}.

@@

Let $F$ be a probability distribution on $\mathbf{C}$ and
assume without loss of generality that $\E f(X) = 0$. Let 
\begin{equation}
\begin{split}
f_k^*, \mu^* \coloneqq \argmin_{f_1,\ldots, f_s, \mu} \Bigl\{ 
             \E ( f(X) &- \sum_{k=1}^s f_k(X_k) -\mu )^2 \\
         &\,:\, \E f_k(X_k) = 0 \Bigr\}.
\end{split}
\end{equation}
We say that $f$ is \emph{additively faithful} under $F$ in case $f^*_k = 0$ iff $f$ does not depend on coordinate $k$. 
\end{definition}
% We can define the support $\trm{supp}(f) \coloneqq \{ k \,:\,
% \trm{$k$ is relevant to $f$}\}$. Let $f^* = \sum_{k=1}^s$, then $f$
% is additively faith if $\trm{supp}(f) = \trm{supp}(f^*)$.

@@

We give the full proof in Section~\ref{sec:faithful_proof} of the
Appendix, but pause here to provide some intuition. From
Lemma~\ref{lem:general_int_reduction}, we know that the
additive approximation zeroes out $k$ when, fixing $x_k$, every
``slice'' of $f$ integrates to zero. We prove
Theorem~\ref{thm:convex_faithful} by showing that ``slices'' of convex
functions that integrate to zero cannot be ``glued'' together while
still maintaining convexity.

Theorem~\ref{thm:convex_faithful} plays an important role in our
sparsistency analysis, where we show that the additive
approximation is variable selection consistent (or ``sparsistent''), even when the true function is not
additive.

\begin{remark}
  We assume twice differentiability in
  Theorem~\ref{thm:convex_faithful} to simplify the proof. We believe
  this smoothness condition is not necessary because every non-smooth
  convex function can be approximated arbitrarily well by a smooth
  one.  
\end{remark}
\begin{remark}
Without restrictions on the distribution, a convex
  function may not be additively faithful. Intuitively, an arbitrarily shaped
  density $p$
  may ``undo'' the convexity of $f$ so that the product
  $p(\mathbf{x}) \, f(\mathbf{x})$ resembles an egg carton or a
  tilting slope.  With appropriate conditions on the density $p$,
  however, it is possible to relax the independence assumption.  We leave this to
  future work.
\end{remark}

% DO NOT CHANGE; RefTex variables -minx

%%% Local Variables: ***
%%% mode:latex ***
%%% TeX-master: "paper.tex" ***
%%% End: ***

### 2013-11-01 tex/scam/appendix.tex
journal

\section*{Appendix}

@@

We start with a lemma similar to Lemma~\ref{lem:general_int_reduction} and whose proofs are identical.

\begin{lemma}
\label{lem:int_reduction}
Let $F$ be a product distribution on $C=[0,1]^s$ with density function $p$. Let $f: C \rightarrow \R$ be a convex function. Suppose that $\E f(X) = 0$.

@@

We first show that the optimal $\mu^* = \E f(X)$ for any $f_1, ..., f_k$ such that $\E f_k(X_k) = 0$. This follows from the stationarity condition, which states that $\mu^* = \E[ f(X) - \sum_k f_k(X_k)] = \E[ f(X) ]$.\\

@@

Now, we verify that as a function of $x_k$, $\E[ f(X) | x_k] - \E f(X)$ has mean zero and is convex. The former is clearly true; the latter is true because for every $\mathbf{x}_{-k}$, $f(x_k, \mathbf{x}_{-k})$ is a convex function with respect to $x_k$ and therefore, $\int_{\mathbf{x}_{-k}} p(\mathbf{x}_{-k}) f(x_k, \mathbf{x}_{-k}) d \mathbf{x}_{-k}$ is still convex.

@@

@@

 \begin{proposition}
Let $F$ be a product distribution supported on $C=[0,1]^s$ with positive density $p > 0$. Let $f: C \rightarrow \R$ be a convex function, twice differentiable.

Let $f^*_1,...,f^*_s \coloneqq \arg\min \{ \E | f(X) -\E f(X) - \sum_k f_k(X_k) |^2 \,:\, \forall k, f_k \trm{ convex }, \E f_k(X_k) = 0\}$.

@@

@@

We let $\langle f, g \rangle_n \coloneqq \frac{1}{n} \sum_{i=1}^n f(X^{(i)}) g(X^{(i)})$. It then follows that:
\begin{packed_enum}
\item $\| f + g \|_n^2 = \|f\|_n^2 + 2 \langle f, g \rangle_n + \|g \|_n^2$
\item $\langle f, g \rangle_n \leq \|f\|_n \| g \_n$
\end{packed_enum}

@@

First, let $\delta = \frac{1}{n}$. Suppose $n$ is large enough such that $ \frac{1}{cn} \log Cn < 1$. Then, we have, with probability at least $1-\frac{1}{n}$,
\begin{align*}
 \frac{1}{n} \sum_{i=1}^n w_i^2 &\leq c\sigma^2 (1+\sqrt{\frac{1}{cn} \log Cn}) \\
		&\leq 2 c \sigma^2
 \end{align*}

>>>>

\section{Appendix}

@@

We start with a proof of Lemma~\ref{lem:general_int_reduction}. We restate the lemma here for convenience.

\begin{lemma}
Let $F$ be a product distribution on $C=[0,1]^s$ with a positive density function $p$. Let $f: C \rightarrow \R$ be an integrable function.

Let $\ds f^*_1,...,f^*_s, \mu^* \coloneqq \arg\min \{ \E | f(X) - \sum_{k=1}^s f_k(X_k) - \mu |^2 \,:\, \forall k, \E f_k(X_k) = 0\}$

@@

We first show that the optimal $\mu^* = \E f(X)$ for any $f_1, ..., f_k$ such that $\E f_k(X_k) = 0$. This follows from the stationarity condition, which states that $\mu^* = \E[ f(X) - \sum_k f_k(X_k)] = \E[ f(X) ]$. \\

@@

Now, we note that as a function of $x_k$, $\E[ f(X) | x_k] - \E f(X)$ has mean zero and we thus finish the proof.

@@

The next lemma adapts Lemma~\ref{lem:general_int_reduction} to convex functions.

\begin{lemma}
\label{lem:int_reduction}
Let $F$ be a product distribution on $C=[0,1]^s$ with positive density function $p$. Let $f: C \rightarrow \R$ be a convex function.

\begin{proof}
The proof is identical to that of Lemma~\ref{lem:general_int_reduction}. We need only verify that $f^*_k(x_k) = \E[ f(X) \given x_k] - \E f(X)$ is convex.

This is true because for every $\mathbf{x}_{-k}$, $f(x_k, \mathbf{x}_{-k})$ is a convex function with respect to $x_k$ and therefore, $\int_{\mathbf{x}_{-k}} p(\mathbf{x}_{-k}) f(x_k, \mathbf{x}_{-k}) d \mathbf{x}_{-k}$ is still convex. 
\end{proof}

@@

\begin{proposition}
\label{prop:faithful_condition}
Let $F$ be a product distribution supported on $C=[0,1]^s$ with positive density $p > 0$. Let $f: C \rightarrow \R$ be a convex function, twice differentiable.

Let $f^*_1,...,f^*_s \coloneqq \arg\min \{ \E | f(X) -\E f(X) - \sum_k f_k(X_k) |^2 \,:\, \forall k, f_k \trm{ convex }, \E f_k(X_k) = 0 \}$.

@@

\textbf{Note:} the symbols $c,C$ represent absolute constants. We will often abuse notation and ``absorb'' new absolute constants into $c, C$; the actual value of $c, C$ could thus vary from line to line.

@@

We let $\langle f, g \rangle_n \coloneqq \frac{1}{n} \sum_{i=1}^n f(X^{(i)}) g(X^{(i)})$. It then follows that:
\begin{packed_enum}
\item $\| f + g \|_n^2 = \|f\|_n^2 + 2 \langle f, g \rangle_n + \|g \|_n^2$
\item $\langle f, g \rangle_n \leq \|f\|_n \| g \|_n$
\end{packed_enum}

@@

First, let $\delta = \frac{1}{n}$. Suppose $n$ is large enough such that $ \frac{1}{cn} \log Cn < 1$. Then, we have, with probability at least $1-\frac{1}{n}$,
\begin{align*}
 \frac{1}{n} \sum_{i=1}^n w_i^2 &\leq c\sigma^2 \Big(1+\sqrt{\frac{1}{cn} \log Cn}\Big) \\
		&\leq 2 c \sigma^2
 \end{align*}

### 2013-11-01 tex/scam/experiments.tex
journal

We next use the Boston housing data rather than simulated data. This data set
contains 13 covariates, 506 samples and one response variable
indicating housing values in suburbs of Boston. The data and detailed description
can be found on the UCI Machine Learning Repository website~\url{http://archive.ics.uci.edu/ml/datasets/Housing}.

@@

\begin{figure*}[!t]
\begin{center}
\begin{tabular}{cccc}
\hskip-20pt
\includegraphics[width=.29\textwidth]{../figs/Curve1} &
\hskip-20pt
\includegraphics[width=.29\textwidth]{../figs/Q} &
\hskip-20pt
\includegraphics[width=.29\textwidth]{../figs/Curve2} &
\hskip-20pt
\includegraphics[width=.29\textwidth]{../figs/Curve3}  \\
\hskip-10pt (a) additive model & 
\hskip-10pt (b) four $\bds{Q}$ matrices &
\hskip-10pt (c) non-additive models & 
\hskip-10pt (d) correlated design
\end{tabular}
\end{center}
\caption{Support recovery results where the additive assumption is
  correct (a), incorrect (b), (c), and with correlated design (d).}\label{Support}
\vskip10pt

\begin{center}
\begin{tabular}{ccc}
\hskip-10pt
  \includegraphics[width=.35\textwidth]{../figs/Additive} &
\hskip-25pt
  \includegraphics[width=.35\textwidth]{../figs/Additive1} &
\hskip-25pt
  \includegraphics[width=.35\textwidth]{../figs/LASSO} 
\\
\hskip-10pt 
SCAM $\|\beta_k\|_\infty$ paths & 
\hskip-25pt 
SCAM $\text{mean}(\beta_k)$ paths &
\hskip-25pt
LASSO $\beta_k$ paths
\end{tabular}
\begin{tabular}{cc}
  \includegraphics[width=.37\textwidth]{../figs/MSE} &
  \includegraphics[width=.43\textwidth]{../figs/Convex}
\\
predictive MSE & inferred functions from SCAM
\end{tabular}
\end{center}
\caption{Results on Boston housing data, showing regularization paths,
 MSE and fitted functions.}\label{Boston}
\end{figure*}

>>>>

We next use the Boston housing data rather than simulated data. This data set
contains 13 covariates, 506 samples and one response variable
indicating housing values in suburbs of Boston. The data and detailed description
can be found on the UCI Machine Learning Repository website~\footnote{\texttt{http://archive.ics.uci.edu/ml/datasets/Housing}}.

@@

\begin{figure*}[!t]
\begin{center}
\begin{tabular}{cccc}
\hskip-20pt
\includegraphics[width=.29\textwidth]{figs/Curve1} &
\hskip-20pt
\includegraphics[width=.29\textwidth]{figs/Q} &
\hskip-20pt
\includegraphics[width=.29\textwidth]{figs/Curve2} &
\hskip-20pt
\includegraphics[width=.29\textwidth]{figs/Curve3}  \\
\hskip-10pt (a) additive model & 
\hskip-10pt (b) four $\bds{Q}$ matrices &
\hskip-10pt (c) non-additive models & 
\hskip-10pt (d) correlated design
\end{tabular}
\end{center}
\caption{Support recovery results where the additive assumption is
  correct (a), incorrect (b), (c), and with correlated design (d).}\label{Support}
\vskip10pt

\begin{center}
\begin{tabular}{ccc}
\hskip-10pt
  \includegraphics[width=.35\textwidth]{figs/Additive} &
\hskip-25pt
  \includegraphics[width=.35\textwidth]{figs/Additive1} &
\hskip-25pt
  \includegraphics[width=.35\textwidth]{figs/LASSO} 
\\
\hskip-10pt 
SCAM $\|\beta_k\|_\infty$ paths & 
\hskip-25pt 
SCAM $\text{mean}(\beta_k)$ paths &
\hskip-25pt
LASSO $\beta_k$ paths
\end{tabular}
\begin{tabular}{cc}
  \includegraphics[width=.37\textwidth]{figs/MSE} &
  \includegraphics[width=.43\textwidth]{figs/Convex}
\\
predictive MSE & inferred functions from SCAM
\end{tabular}
\end{center}
\caption{Results on Boston housing data, showing regularization paths,
 MSE and fitted functions.}\label{Boston}
\end{figure*}

% DO NOT CHANGE; RefTex variables -minx

%%% Local Variables: ***
%%% mode:latex ***
%%% TeX-master: "paper.tex" ***
%%% End: ***

### 2013-11-01 tex/scam/icml14/additive_faithful_icml.tex
updated proof of additive faithfulness
journal

\begin{lemma}
\label{lem:general_int_reduction}
Let $F$ be a product distribution on $\mathbf{C}=[0,1]^s$ with density function $p$ which is positive on $\mathbf{C}$. Let
$X=(X_1,...,X_s) \sim F$. Let $f: \mathbf{C} \rightarrow \R$ be an
integrable function and suppose, without loss of generality, that $\E f(X) = 0$.
Let 
\begin{equation}
f^* \coloneqq \argmin_{f_1,\ldots, f_s} \Bigl\{ \E \bigl( f(X) -
\sum_{k=1}^s f_k(X_k) \bigr)^2 \,:\, \E f_k(X_k) = 0\Bigr\}.
\end{equation}
Then $f^*_k(x_k) = \E[ f(X) \given x_k]$ is unique.
\end{lemma}

@@

Let $F$ be a probability distribution on $\mathbf{C}$ and
assume without loss of generality that $\E f(X) = 0$. Let 
\begin{equation}
f^* \coloneqq \argmin_{f_1,\ldots, f_s} \Bigl\{ \E ( f(X) - \sum_{k=1}^s
f_k(X_k) )^2 \,:\, \E f_k(X_k) = 0 \Bigr\}.
\end{equation}
We say that $f$ is \emph{additively faithful} under $F$ in case $f^*_k = 0$ iff $f$ does not depend on coordinate $k$. 
\end{definition}
% We can define the support $\trm{supp}(f) \coloneqq \{ k \,:\,
% \trm{$k$ is relevant to $f$}\}$. Let $f^* = \sum_{k=1}^s$, then $f$
% is additively faith if $\trm{supp}(f) = \trm{supp}(f^*)$.

@@

\begin{remark}
  We assume twice differentiability in
  Theorem~\ref{thm:convex_faithful} to simplify the proof. We believe
  this smoothness condition is not necessary because every non-smooth
  convex function can be approximated arbitrarily well by a smooth
  one.  Without restrictions on the distribution, a convex
  function may not be additively faithful. Intuitively, an arbitrarily shaped
  density $p$
  may ``undo'' the convexity of $f$ so that the product
  $p(\mathbf{x}) \, f(\mathbf{x})$ resembles an egg carton or a
  tilting slope.  With appropriate conditions on the density $p$,
  however, it is possible to relax the independence assumption.  We leave this to
  future work.
\end{remark}

>>>>

\begin{lemma}
\label{lem:general_int_reduction}
Let $F$ be a product distribution on $\mathbf{C}=[0,1]^s$ with density function $p$ which is positive on $\mathbf{C}$. Let
$X=(X_1,...,X_s) \sim F$. Let $f: \mathbf{C} \rightarrow \R$ be an
integrable function.
Let 
\begin{equation}
\begin{split}
f_k^*, \mu^* \coloneqq \argmin_{f_1,\ldots, f_s, \mu} \Bigl\{ \E \bigl( f(X) & -
 \sum_{k=1}^s f_k(X_k) -\mu \bigr)^2 \\
       & \,:\, \E f_k(X_k) = 0\Bigr\}.
\end{split}
\end{equation}
Then $\mu^* = \E f(X)$ and $f^*_k(x_k) = \E[ f(X) \given x_k] - \E f(X)$ and this solution is unique.
\end{lemma}

@@

Let $F$ be a probability distribution on $\mathbf{C}$ and
assume without loss of generality that $\E f(X) = 0$. Let 
\begin{equation}
\begin{split}
f_k^*, \mu^* \coloneqq \argmin_{f_1,\ldots, f_s, \mu} \Bigl\{ 
             \E ( f(X) &- \sum_{k=1}^s f_k(X_k) -\mu )^2 \\
         &\,:\, \E f_k(X_k) = 0 \Bigr\}.
\end{split}
\end{equation}
We say that $f$ is \emph{additively faithful} under $F$ in case $f^*_k = 0$ iff $f$ does not depend on coordinate $k$. 
\end{definition}
% We can define the support $\trm{supp}(f) \coloneqq \{ k \,:\,
% \trm{$k$ is relevant to $f$}\}$. Let $f^* = \sum_{k=1}^s$, then $f$
% is additively faith if $\trm{supp}(f) = \trm{supp}(f^*)$.

@@

\begin{remark}
  We assume twice differentiability in
  Theorem~\ref{thm:convex_faithful} to simplify the proof. We believe
  this smoothness condition is not necessary because every non-smooth
  convex function can be approximated arbitrarily well by a smooth
  one.  
\end{remark}
\begin{remark}
Without restrictions on the distribution, a convex
  function may not be additively faithful. Intuitively, an arbitrarily shaped
  density $p$
  may ``undo'' the convexity of $f$ so that the product
  $p(\mathbf{x}) \, f(\mathbf{x})$ resembles an egg carton or a
  tilting slope.  With appropriate conditions on the density $p$,
  however, it is possible to relax the independence assumption.  We leave this to
  future work.
\end{remark}

### 2013-11-01 tex/scam/icml14/introduction_icml.tex
journal

The general sparse nonparametric regression problem is considered in
\cite{lafferty2008rodeo}, where it is shown that computationally
efficient, near minimax-optimal estimation is possible, but in ambient
dimensions that scale only as $p = O(\log n)$.  This is in stark
contrast to the exponential scaling $p = O\bigl(e^{n^c}\bigr)$ enjoyed
by sparse linear models \cite{Wain:09a}.  Sparse additive models
\cite{Ravikumar:09} provide a practical alternative to fully
nonparametric function estimation.  But the additive assumption is
limited.  In particular, the natural idea of first selecting the
single variable effects, then the pairwise effects, and so on, does
not in general lead to consistent variable selection.  In other words,
the general nonparametric model is not additively faithful.
Remarkably, the additional assumption of convexity does lead to
additive faithfulness, as we show here.  In addition, we show that
exponential scaling is achievable for sparse convex additive models.
Thus, the geometric convexity constraint is quite different from the
smoothness constraints imposed in traditional nonparametric
regression.

@@

Estimation of convex functions arises naturally in several
applications.  Examples include geometric programming \cite{Boyd04},
computed tomography \cite{Prince:90}, target reconstruction
\cite{Lele:92}, image analysis \cite{Golden:06} and circuit design
\cite{Hannah:12}.  Other applications include queuing theory
\cite{Chen:01} and economics, where it is of interest to estimate
concave utility functions \cite{Pratt:68}.  See \cite{Lim:12} for
other applications.  Beyond cases where the assumption of convexity is
natural, we offer that the convexity assumption is attractive as a
tractable, nonparamametric relaxation of the linear model.  In
addition to the lack of tuning parameters, other than the
regularization parameter $\lambda$ to control the level of sparsity,
the global convexity assumption leads to effective, scalable algorithms.  We
demonstrate use of our approach on experiments with standard
regression data sets, in a comparison with sparse linear models
(lasso).

@@

\textbf{Notation.} If
$\mathbf{x}$ is a vector, we use $\mathbf{x}_{-k}$ to denote the
vector with the $k$-th coordinate removed. If $\mathbf{v} \in \R^n$, then
$v_{(1)}$ denotes the smallest coordinate of $\mathbf{v}$ in
magnitude, and $v_{(j)}$ denotes the $j$-th smallest; $\mathbf{1}_n \in \R^n$
is the all ones vector. If $X \in \R^p$ and $S \subset
\{1,...,p\}$, then $X_S$ is the subvector of $X$ restricted to
the coordinates in $S$. Given $n$ samples $X^{(1)},...,X^{(n)}$, we use
$\bar{X}$ to denote the empirical average.

>>>>

The general sparse nonparametric regression problem is considered in \cite{lafferty2008rodeo}, where it is shown that computationally efficient, near minimax-optimal estimation is possible, but in ambient
dimensions that scale only as $p = O(\log n)$ instead of $p=O\bigl(e^{n^c}\bigr)$ as enjoyed by sparse linear models. \citet{dalalyan:12} do achieve exponential scaling $p=O(e^n)$ under certain Fourier smoothness conditions but they show that variable selection is still hard in that the number of relevant variables $s$ must be less than $\log n$.

% The general sparse nonparametric regression problem is considered in
% \cite{lafferty2008rodeo}, where it is shown that computationally
% efficient, near minimax-optimal estimation is possible, but in ambient
% dimensions that scale only as $p = O(\log n)$.  This is in stark
% contrast to the exponential scaling $p = O\bigl(e^{n^c}\bigr)$ enjoyed
% by sparse linear models \cite{Wain:09a}.

Approximating the regression function by a sum of one-dimensional functions, known as sparse additive models, \cite{Ravikumar:09} is a practical alternative to fully
nonparametric function estimation.  But the additive assumption is
limited.  In particular, the natural idea of first selecting the
single variable effects, then the pairwise effects, and so on, does
not in general lead to consistent variable selection.  In other words,
the general nonparametric model is not additively faithful.
Remarkably, the additional assumption of convexity does lead to
consistent variable selection, as we show here. In addition, we show that
the scaling $p = O(\log n)$ and $n = O\big(\textrm{poly}(s)\big)$ is achievable for sparse convex additive models. Thus, the geometric convexity constraint is quite different from the
smoothness constraints imposed in traditional nonparametric
regression.

@@

Estimation of convex functions arises naturally in several
applications.  Examples include geometric programming \cite{Boyd04},
computed tomography \cite{Prince:90}, target reconstruction
\cite{Lele:92}, image analysis \cite{Golden:06} and circuit design
\cite{Hannah:12}.  Other applications include queuing theory
\cite{Chen:01} and economics, where it is of interest to estimate
concave utility functions \cite{Pratt:68}.  See \cite{Lim:12} for
other applications.

Beyond cases where the assumption of convexity is
natural, we offer that the convexity assumption is attractive as a
tractable, nonparamametric relaxation of the linear model.  In
addition to the lack of tuning parameters, other than the
regularization parameter $\lambda$ to control the level of sparsity,
the global convexity assumption leads to effective, scalable algorithms.  We
demonstrate use of our approach on experiments with standard
regression data sets, in a comparison with sparse linear models
(lasso).

@@
